---
title: "Data Wrangling Project 1"
output: pdf_document
---

## Objective 
The objective of this project is to deal with missing values in a toy dataset showing product purchases from an electronics store as a part of Springboard's Foundations of Data Science course.

## Setup
The first step is to load the required libraries and the dataset in R.
Load libraries:

```{r}
library (tidyr)
library (dplyr)
library(stringr)
library(psych)
```

Import .csv file as  a data frame:

```{r}
raw_data <- read.csv("/Users/tanvidesai/Documents/2016 Courses/Foundations of Data Science/Data Wrangling Project 1/refine_original.csv", 
         head = TRUE, sep = ",")
tbl_df(raw_data)
View(raw_data)
```

## Brand names
The 'company' column has four brands: Philips, Akzo, Van Houten and Unilever. There are many different spellings and capitaliations for these. So we will clean up the brand names by transforming them.

```{r}
raw_data$company <- tolower(raw_data$company)
raw_data$company <- str_replace(raw_data$company, "f", "ph")
raw_data$company <- str_replace(raw_data$company, "^ph[a-z]+ps", "philips")
raw_data$company <- str_replace(raw_data$company, "^ak[a-z0-9 ]+", "akzo")
raw_data$company <- str_replace(raw_data$company,"unilver","unilever")
```

## Product code and number
The product code and number are combined into a single column and we would like to separate them into different columns.

```{r}
raw_data <- separate(raw_data, Product.code...number, 
                     c("product_code", "product_number"), sep = "-")
```

## Product categories
The product codes represent the following product categories:
- p = Smartphone
- v = TV
- x = Laptop
- q = Tablet
We will add a column with the product category for each record.

```{r}
categorize_product <- function(product_code) {
  if (product_code == "p") {
  return ("Smartphone") 
  } else if (product_code == "v") {
    return ("TV") 
  } else if (product_code == "x") {
    return ("Laptop") 
  } else if (product_code == "q") {
    return("Tablet") 
  }}
```

## Address
We will combine the separated address components into a single address column that can be easily geocoded.

```{r}
raw_data <- unite(raw_data, "full_address", address, city, country, sep = ",")
```

## Dummy variables for company and product category
We will create binary dummy variables for each set of values in company and product category.
1) Company
```{r}
dummy_var <- dummy.code(raw_data$company)
new_dataset <- data.frame(raw_data, dummy_var)
colnames(new_dataset)[7:10] <- paste("company_", colnames(new_dataset[ c(7:10)]), sep = "")
```

2) Product category
```{r}
dummy_var1 <- dummy.code(new_dataset$product_category)
final_dataset <- data.frame(new_dataset, dummy_var1)
colnames(final_dataset)[11:14] <- paste("product_", colnames(final_dataset[ c(11:14)]), sep = "")

View(final_dataset)
```

## Output
Finally, we will export the new dataset as .csv file.
```{r}
write.csv(final_dataset, "/Users/tanvidesai/Documents/2016 Courses/Foundations of Data Science/Data Wrangling Project 1/refine_cleaned.csv")
```

